\section{Experimental results}

\subsection{Setup}
\label{sec:experiments}
\input{tables/main_results}
We evaluate \methodname on Llama-3-8B~\cite{grattafiori2024llama3herdmodels}.
We compare against the uncompressed FP16 model and the most accurate SVD-style baseline, EigenAttention~\cite{saxena2024eigenattention}, under matched KV cache budgets.

We report results on both language modeling and zero-shot multiple-choice benchmarks.
For language modeling, we measure perplexity on WikiText-2~\cite{wikitext} and C4~\cite{c4} datasets.
For zero-shot evaluation, we use the LM Evaluation Harness framework~\cite{eval-harness} and report accuracy on HellaSwag~\cite{hellaswag}, PIQA~\cite{piqa}, and MMLU~\cite{mmlu1, mmlu2}. We run all experiments on two NVIDIA RTX A5000 GPUs, with 24\,GB each.

% \paragraph{KV cache budgets.}
% We compress the KV cache along the per-head feature dimension by storing low-rank projections (Sec.~\ref{sec:background}).
% We sweep compression levels by varying the per-layer error budget $\epsilon_\ell$ and selecting layer-wise ranks $(r_K, r_V)$ such that $\Delta_\ell(r_K, r_V; \theta) \le \epsilon_\ell$; different budgets induce different KV cache compression ratios.

Unless stated otherwise, ranks are allocated per layer %using the error surfaces computed on the calibration set from Sec.~\ref{sec:method}, 
following the Weighted Pareto policy of Sec.~\ref{sec:rank_selection}.
%We show the advantage of this rank-allocation policy in Sec.~\ref{sec:results_policies}.

\paragraph{Calibration data and training.}
All projection bases and rank profiles are obtained post-training from a held-out calibration set of 512 sequences sampled from WikiText-2~\cite{wikitext}.
Due to GPU memory constraints, we use a sequence length of 2048 for both calibration and all evaluations, unless stated otherwise.
For \methodname, we train the basis predictor independently for 5 down-projection ranks uniformly spaced between 50\% and 90\% of the per-head dimension $d_h$. We train for up to 50 epochs with early stopping (patience 5, delta $10^{-6}$).
We use batch size 1 for keys and batch size 4 for values; we found value bases to be more numerically unstable and therefore use the largest stable batch size supported by our hardware.
We optimize with AdamW (learning rate $5\cdot 10^{-3}$, weight decay $10^{-4}$) and a cosine annealing schedule with $T_{\max}$ set to the total number of training steps. We do not fine-tune on downstream tasks.
For a fair comparison, we re-run EigenAttention in our setup, matching calibration data, sequence length, and evaluation protocol, and report all baselines under these identical conditions.

