\section{Related Work} \label{sec:related_work}

\paragraph{Improving Inference Efficiency} Many works aim to reduce the inference cost of neural networks, typically via quantization and pruning. Quantization reduces operation costs by lowering the precision of weights, while pruning removes partial connections from the network to reduce the number of operations. We refer to \citet{Gholami2021QuantSurvey} and \citet{liu2025low} for a comprehensive survey on quantization, and to \citet{cheng2024pruningsurvey} for a comprehensive survey on pruning. These methods motivate the study of Boolean networks, which replace floating-point operations with Boolean operations, significantly reducing operation cost and connection density simultaneously.

\paragraph{Learning Boolean Structures} We review prior works on learning three different Boolean structures: Boolean networks, binary decision diagrams (BDDs), and lookup-tables (LUTs). \citet{petersen2022deep} first proposed differentiable learning of Boolean networks. Many later works improved the learning of Boolean networks: \citet{ruttgers2025light} reduced the parameter overhead during training; \citet{bacellar2024differentiable} and \citet{fojcik2025lilogic} introduced learnable connections between layers; \citet{petersen2024convolutional} designed convolutional Boolean networks; \citet{buhrer2025recurrent} extended Boolean networks to recurrent architectures. For BDDs, \citet{Shati2023SATBDD} and \citet{Florio2023SATBDD} formulated BDD learning as a SAT problem; \citet{Zantedeschi2021ArgminTrees} trained discrete decision trees via argmin differentiation; \citet{Qiu2025OBDDNet} proposed to parameterize ordered BDDs with continuous encodings, and then read off discrete BDDs after training. For LUTs, \citet{Umuroglu2020LogicNetsFCCM} reduced sparse quantized networks into LUT networks; \citet{Wang2019LUTNet} trained LUT networks via Lagrange-interpolating polynomial relaxations; \citet{Andronic2025NeuraLUTAssembleFCCM} proposed a hardware-aware training and assembly strategy for LUT networks; \citet{Cassidy2025ReducedLUT} performed post-training LUT compression; \citet{Weng2025AmigoLUT} assembles multiple small LUT subnetworks to form a larger LUT network.
