\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\drew}[1]{\textcolor{blue}{[Dr: #1]}}


\documentclass[preprint]{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subcaption} % figures
\usepackage{booktabs} % for professional tables
\usepackage{bbm}

\usepackage{tikz}

\usepackage{xcolor}

\usepackage{hyperref}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{stmaryrd,scalerel}

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2025} with \usepackage[nohyperref]{icml2025} above.

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2026}

% If accepted, instead use the following line for the camera-ready submission:
% \usepackage{icml2026}

% For theorems and such
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}

\usepackage{pifont}% http://ctan.org/pkg/pifont

\newcommand{\cmark}{\Large\textcolor{green!80!black}{\ding{51}}}
\newcommand{\xmark}{\Large\textcolor{red}{\ding{55}}}
\newcommand{\mygreen}[1]{{\color{green!60!black}#1}}


% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
\usepackage[disable,textsize=tiny]{todonotes}
%\usepackage[textsize=tiny]{todonotes}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
% \icmltitlerunning{Generalized Conformal Risk Control for Safe Policy Improvement}
\icmltitlerunning{Conformal Policy Control}

\begin{document}

\twocolumn[
\icmltitle{Conformal Policy Control}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2025
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}\icmlauthor{Anonymous}{anon}\end{icmlauthorlist}

\icmlaffiliation{anon}{Anonymous Institution}




\icmlcorrespondingauthor{Anonymous}{anonymous@anonymous.invalid}


% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{machine-learning, conformal-risk-control, model-based-optimization, safe-exploration, conservative-optimization}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
% \printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.


\begin{abstract}
    An agent must try new behaviors to explore and improve. In high-stakes environments, an agent that violates safety constraints may cause harm and must be taken offline, curtailing any future interaction. Imitating old behavior is safe, but excessive conservatism discourages exploration. How much behavior change is too much? We show how to use any safe reference policy as a probabilistic regulator for any optimized but untested policy. Conformal calibration on data from the safe policy determines how aggressively the new policy can act, while provably enforcing the user's declared risk tolerance. Unlike conservative optimization methods, we do not assume the user has identified the correct model class nor tuned any hyperparameters. Unlike previous conformal methods, our theory provides finite-sample guarantees even for non-monotonic bounded loss functions. Our experiments on applications ranging from natural language question answering to biomolecular engineering show that safe exploration is not only possible from the first moment of deployment, but can also improve performance.
\end{abstract}


\input{files/1_Introduction}

\vspace{-0.25cm}

\input{files/2_Background}


\input{files/3_RelatedWork}


\input{files/4_MethodsTheory}


\input{files/5_Experiments}


\input{files/6_Conclusion}

\clearpage

\section*{Code Availability}

Code to reproduce the results is available at \url{https://github.com/samuelstanton/conformal-policy-control}.

\input{files/7_BroaderImpact}

\bibliography{references}
\bibliographystyle{icml2026}

\newpage
\appendix
\onecolumn

\begin{center}
{\Large\bfseries Supplementary Material}
\end{center}
\vspace{1em}

\section*{Table of Contents}
\newcommand{\appentry}[3]{\noindent\textbf{#1}\hspace{1em}\hyperref[#2]{#3}\dotfill\pageref{#2}\par\vspace{0.3em}}
\appentry{A}{app:extended_related_work}{Extended Related Work}
\appentry{B}{app:proofs}{Proofs}
\appentry{C}{app:detailed_methodology}{Implementation Details}
\appentry{D}{app:medical_qa}{Medical QA Experiments}
\appentry{E}{app:active_learning_details}{Tabular Active Learning Experiments}
\appentry{F}{app:sequence_opt}{Sequence Optimization Experiments}

\input{files/A_ExtendedRelatedWork}

\input{files/B_Proofs}

\input{files/C_Method_Details}

\input{files/D_MedicalQA}

\input{files/E_TabularActiveLearning}

\input{files/F_SequenceOptimization}

\end{document}
