\def\isarxiv{1}

\ifdefined\isarxiv
\documentclass[12pt]{article}
\usepackage[margin=1.25in]{geometry} 
\usepackage{times}
\usepackage{algpseudocode}
\usepackage{hyperref} 
\else
\documentclass{article}%[nonacm,anonymous,acmsmall,pdftex]{acmart}
\usepackage{hyperref}       % hyperlinks
\usepackage{multicol,enumitem}
% \usepackage{enumitem}
% \setlist{nosep}

\usepackage{icml2026}
% \usepackage{times}
% \usepackage{algpseudocode}
% \usepackage{algorithmic}
% \usepackage[noend]{algpseudocode} % Load algorithmic package first
% \usepackage{algorithm} % Load algorithm package
\fi


\usepackage{natbib}
\usepackage{url}   

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

% \usepackage{xcolor}         % colors
%\usepackage{tablefootnote}

%****************************************************
% \usepackage{cleveref}
\usepackage{cleveref}
\usepackage{comment}
\usepackage{bbm}
%\usepackage[ruled, vlined]{algorithm2e}
\usepackage{textcomp}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{wrapfig}

%****************************************************
%The following are not in original neurips template
% \usepackage[utf8]{inputenc} % allow utf-8 input
% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
% \usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{multirow}
% \usepackage{algorithm}
%\usepackage{algorithmic}
% \usepackage{algorithm}
%\usepackage{subfig}
\usepackage{color}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{grffile}
%\usepackage{natbib}
%\usepackage[pdf]{pstricks}
\usepackage{wrapfig,epsfig}
%\usepackage{psfrag}
% \usepackage{epstopdf}
% \usepackage{epstopdf}
% \usepackage{algpseudocode}
%\usepackage{scrextend}
%\usepackage[hidelinks,pdfencoding=auto,psdextra]{hyperref} %%% commet for arxiv
% \usepackage[T1]{fontenc}
\usepackage{bbm}
\usepackage{comment}
% \usepackage{caption}
\usepackage{subcaption} % don't add this packge
% \usepackage{dsfont}
\usepackage{tikz}
\usepackage{colortbl}
%\usepackage{ftnxtra}
%\usepackage{tablefootnote}
%\usetikzlibrary{graphs,graphdrawing,arrows}

\usepackage{nicematrix}
\usepackage{makecell}

\usepackage{pifont} % render ding marks
% \usepackage{arydshln} % add dashed lines in tables
%****************************************************


%\SetKwInOut{Parameter}{parameter}

\usepackage{pgfplots}
\pgfplotsset{compat=1.8}
\tikzset{elegant/.style={smooth,thick,samples=500,magenta}}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{example}[theorem]{Example}
\newtheorem{problem}[theorem]{Problem}

\newtheorem{fact}[theorem]{Fact}
\newtheorem{claim}[theorem]{Claim}
\crefname{assumption}{Assumption}{Assumptions}


\theoremstyle{plain}
\newtheorem{thm}{Theorem} % definition numbers are dependent on theorem numbers
\newtheorem*{thm*}{Theorem}
%\newtheorem{lemma}{Lemma}


\theoremstyle{plain}
\newtheorem{prop}{Proposition}
%\newtheorem{fact}{Fact}
% \newtheorem{claim}{Claim}
% \newtheorem{corollary}{Corollary}


%\theoremstyle{definition}
\newtheorem{defn}{Definition} % definition numbers are dependent on theorem numbers

%\theoremstyle{remark}
%\newtheorem{remark}{Remark} 



%****************************************************

%\input{math_commands}
\input{symbols}


\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\blue}[1]{{\color{blue}#1}}

\newcommand{\Baihe}[1]{{\color{mygreen}[Baihe: #1]}}

\newcommand{\todo}[1]{{\color{red}[TODO: #1]}}

\newcommand{\Eric}[1]{{\color{mint}[Eric: #1]}}
%\newcommand{\blue}[1]{{\color{blue}#1}}
\definecolor{b2}{RGB}{51,153,255}
\definecolor{mygreen}{RGB}{80,180,0}
\definecolor{mint}{RGB}{62, 180, 137}

\newcommand{\vct}{\boldsymbol }
\newcommand{\sR}{\mathfrak R}

\title{Towards Anytime-Valid Statistical Watermarking}


\begin{document}

\ifdefined\isarxiv
\author{
Baihe Huang\thanks{
  \texttt{baihe\_huang@berkeley.edu}. University of California, Berkeley. }
  % examples of more authors
  \and
  Eric Xu\thanks{\texttt{erx@berkeley.edu}. University of California, Berkeley. }
  \and
  Kannan Ramchandran
  \thanks{
  \texttt{kannanr@berkeley.edu}. University of California, Berkeley. }
  \and
  Jiantao Jiao\thanks{
  \texttt{jiantao@eecs.berkeley.edu}. University of California, Berkeley. }
  \and
  Michael I. Jordan\thanks{
  \texttt{jordan@cs.berkeley.edu}. University of California, Berkeley. }
}
\date{}
\maketitle

\else
\twocolumn[
  \icmltitle{Towards Anytime-Valid Statistical Watermarking}

  % It is OKAY to include author information, even for blind submissions: the
  % style file will automatically remove it for you unless you've provided
  % the [accepted] option to the icml2026 package.

  % List of affiliations: The first argument should be a (short) identifier you
  % will use later to specify author affiliations Academic affiliations
  % should list Department, University, City, Region, Country Industry
  % affiliations should list Company, City, Region, Country

  % You can specify symbols, otherwise they are numbered in order. Ideally, you
  % should not use this facility. Affiliations will be numbered in order of
  % appearance and this is the preferred way.
  \icmlsetsymbol{equal}{*}

  \begin{icmlauthorlist}\icmlauthor{Anonymous}{anon}\end{icmlauthorlist}

  \icmlaffiliation{anon}{Anonymous Institution}
  
  

  \icmlcorrespondingauthor{Anonymous}{anonymous@anonymous.invalid}
  

  % You may provide any keywords that you find helpful for describing your
  % paper; these are used to populate the "keywords" metadata in the PDF but
  % will not be shown in the document
  \icmlkeywords{Machine Learning, ICML}

  \vskip 0.3in
]

\fi


% \maketitle
\setlength{\abovedisplayskip}{3.2pt}
\setlength{\belowdisplayskip}{3.2pt}
% \printAffiliationsAndNotice{}

\begin{abstract}
The proliferation of Large Language Models (LLMs) necessitates efficient mechanisms to distinguish machine-generated content from human text. While statistical watermarking has emerged as a promising solution, existing methods suffer from two critical limitations: the lack of a principled approach for selecting sampling distributions and the reliance on fixed-horizon hypothesis testing, which precludes valid early stopping. In this paper, we bridge this gap by developing the first e-value-based watermarking framework, \emph{Anchored E-Watermarking}, that unifies optimal sampling with anytime-valid inference. Unlike traditional approaches where optional stopping invalidates Type-I error guarantees, our framework enables valid, anytime-inference by constructing a test supermartingale for the detection process. By leveraging an anchor distribution to approximate the target model, we characterize the optimal e-value with respect to the \emph{worst-case log-growth rate} and derive the optimal expected stopping time. Our theoretical claims are substantiated by simulations and evaluations on established benchmarks, showing that our framework can significantly enhance sample efficiency, reducing the average token budget required for detection by 13-15\% relative to state-of-the-art baselines.
\end{abstract}


\input{intro}
\input{prelim}
\input{problem}
\input{theory}
\input{experiment}
\input{discussions}

\ifdefined\isarxiv
\bibliography{ref}
\bibliographystyle{icml2026}
\newpage
\appendix
\else

\bibliography{ref}
\bibliographystyle{icml2026}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\onecolumn

\fi


\newpage
\appendix

\input{e_proof}
\input{tau_proof}
\input{claims}
\input{add_experiments}

\end{document}
