\section{Proof of \Cref{thm:stopping-time}}\label{sec:tau-proof}

\begin{proof}
We prove the first claim: Let $q^*$ and $p^*$ be the solution of the problem in Eq.~\eqref{eq:log-growth} for the e-value $e$. Define the adversary $\cA^*$ that selects $q^t \equiv q^*$ for all $t \in \Z_+$. Then \Cref{thm:log-growth} implies that 
\begin{align*}
\sup_{\cG} \mathbb{E}_{\mu(\cA^*,\cG)}[\log e(v^t,s^t)] = &~
    \sup_{w \in \cP(p_0,q^*)} \sum_{v,s \in \cV} w(v,s) \cdot \log e(v,s) \\
    \leq &~ J^*.
\end{align*}
Applying \Cref{thm:dynamic-hitting-time}, we have 
\begin{align*}
    \inf_{\cG} \liminf_{\alpha \downarrow 0} \frac{\mathbb{E}_{\mu(\cA^*,\cG)}[\tau_\alpha(e)]}{\log(1/\alpha)} = \frac{1}{\sup_{w \in \cP(p_0,q^*)} \sum_{v,s \in \cV} w(v,s) \cdot \log e(v,s)} \geq \frac{1}{J^*}.
\end{align*}
This establishes the first claim.

For the e-value given by
\begin{align*}
    e^*(v^t,s^t) = \begin{cases}
       \frac{1-\delta/2}{p_0(s^t)}, & v^t = s^t,\\
       \frac{\delta}{2(n-1)p_0(s^t)}, & v^t\neq s^t,
    \end{cases}
\end{align*}
\Cref{thm:log-growth} implies that 
\begin{align*}
    \inf_{q \in \cQ(p_0,\delta)} \sup_{w \in \cP(p_0,q)} &~  \sum_{v,s \in \cV} w(v,s) \cdot \log e^*(v,s) \geq J^*.
\end{align*}
It follows that for any adversary $\cA$, there exists a generator $\cG$ such that
\begin{align*}
    \mathbb{E}_{\mu(\cA,\cG)}[\log e^*(v^t,s^t)] \geq J^* ,\quad \forall t \in \Z_+.
\end{align*}
Applying \Cref{thm:dynamic-hitting-time-lower-bound-drift}, we have for any adversary $\cA$
\begin{align*}
    \inf_{\cG} \liminf_{\alpha \downarrow 0} \frac{\mathbb{E}_{\mu(\cA,\cG)}[\tau_\alpha(e^*)]}{\log(1/\alpha)} \leq \frac{1}{J^*}.
\end{align*}
This establishes the second claim.
\end{proof}

\subsection{Useful results}

\begin{theorem}[Dynamic robust sample complexity with converging drift]
\label{thm:dynamic-hitting-time}
Fix a filtered probability space $(\Omega,\mathcal{F},(\mathcal{F}_t)_{t\geq 0},\mathbb{P})$.
Let $(Y_t)_{t\geq 1}$ be a sequence of integrable random variables adapted to $(\mathcal{F}_t)$, and define the partial sums
\begin{align*}
S_t \;:=\; \sum_{i=1}^t Y_i, \qquad t\geq 1,
\end{align*}
with the convention $S_0:=0$.

Assume the following.

\begin{enumerate}
  \item[(A1)] (\emph{Bounded increments}) There exists a constant $M\in(0,\infty)$ such that
  \begin{align*}
    |Y_t|\;\leq\; M \quad \text{almost surely for all } t\geq 1.
  \end{align*}
  In particular, $Y_t \leq M$ almost surely for all $t$.

  \item[(A2)] (\emph{Positive, converging conditional drift}) There exists a deterministic sequence $(J_t)_{t\geq 1}$ and a constant $J_{\inf}>0$ such that
  \begin{align*}
    \mathbb{E}\!\bigl[\,Y_t \,\bigm|\, \mathcal{F}_{t-1}\,\bigr] \;=\; J_t
    \quad\text{almost surely for all } t\geq 1,
  \end{align*}
  and
  \begin{align*}
    J_{\inf} \;\leq\; J_t \;\leq\; J_{\sup} < \infty
    \quad\text{for all } t\geq 1,
  \end{align*}
  for some finite $J_{\sup}$, and moreover
  \begin{align*}
    J_t \;\longrightarrow\; J_\infty \in (0,\infty)
    \quad\text{as } t\to\infty.
  \end{align*}

  \item[(A3)] (\emph{Stopping rule}) For each threshold $B>0$, define the stopping time
  \begin{align*}
    \tau_B \;:=\; \inf\{\,t\geq 1 : S_t \geq B\,\},
  \end{align*}
  with the usual convention $\inf\emptyset := +\infty$.
\end{enumerate}

Then for every $B>0$ the stopping time $\tau_B$ is integrable, and as $B\to\infty$,
\begin{align*}
  \frac{\mathbb{E}[\tau_B]}{B}
  \;\longrightarrow\; \frac{1}{J_\infty}.
\end{align*}

Equivalently, if we define $B_\alpha := \log(1/\alpha)$ and
\begin{align*}
  \tau_\alpha := \inf\{\,t\geq 1 : S_t \geq B_\alpha\,\}, \qquad \alpha\in(0,1),
\end{align*}
then
\begin{align*}
  \lim_{\alpha\downarrow 0}
  \frac{\mathbb{E}[\tau_\alpha]}{\log(1/\alpha)}
  \;=\; \frac{1}{J_\infty}.
\end{align*}
\end{theorem}

\begin{proof}
We break the proof into several steps. The argument is self-contained and uses only basic properties of conditional expectation and stopping times.

\medskip
\noindent We first establish boundedness of $S_{\tau_B}$ on the event that $\tau_B$ is finite. For each fixed $B>0$, let $\tau_B$ be as in (A3). Because $(Y_t)$ is adapted and the condition $\{S_t\geq B\}$ depends only on $Y_1,\dots,Y_t$, $\tau_B$ is a stopping time with respect to $(\mathcal{F}_t)$.

By definition of $\tau_B$,
\begin{align*}
  S_{\tau_B} \geq B
  \quad\text{on the event } \{\tau_B<\infty\},
\end{align*}

On the event $\{\tau_B<\infty\}$, we have $S_{\tau_B-1} < B$
and $Y_{\tau_B}\leq M$ from the bounded increments assumption (A1). Hence
\begin{align*}
  S_{\tau_B} = S_{\tau_B-1} + Y_{\tau_B}
  < B + M
  \quad\text{on } \{\tau_B<\infty\}.
\end{align*}
Combining the two inequalities gives
\begin{equation}
  B \mathbf{1}_{\{\tau_B<\infty\}}
  \leq
  S_{\tau_B}\mathbf{1}_{\{\tau_B<\infty\}}
  <
  (B+M)\mathbf{1}_{\{\tau_B<\infty\}}.
  \label{eq:overshoot-inequality}
\end{equation}

\medskip


Next, we show that $\tau_B$ is integrable and obtain a crude upper bound on its expectation that will be used later.

For $n\in\mathbb{N}$ define the truncated stopping time
\begin{align*}
  \tau_B^{(n)} := \inf\{\tau_B, n\},
\end{align*}

which is integrable for each fixed $n$. On the one hand,
\begin{align*}
  S_{\tau_B^{(n)}}
  =
  \sum_{t=1}^{\tau_B^{(n)}} Y_t
  = \sum_{t=1}^n Y_t \mathbf{1}_{\{\tau_B\geq t\}},
\end{align*}
because the sum stops at $t=\tau_B$ if $\tau_B\leq n$, and otherwise at $t=n$ if $\tau_B>n$. 

Linearity of expectation yields
\begin{equation}
  \mathbb{E}\bigl[S_{\tau_B^{(n)}}\bigr]
  =
  \sum_{t=1}^n \mathbb{E}\bigl[Y_t \mathbf{1}_{\{\tau_B\geq t\}}\bigr].
  \label{eq:Es-tau-n-expansion}
\end{equation}
This exchange of summation and expectation is justified because $Y_t$ is bounded a.s. for all $t$.

Now we use the conditional drift assumption (A2). Because $Y_t$ is $\mathcal{F}_t$-measurable and $\mathcal{F}_{t-1}$-adapted, and $\{\tau_B\geq t\} = \{\tau_B>t-1\}\in\mathcal{F}_{t-1}$ (by the definition of a stopping time), we have
\begin{align*}
  \mathbb{E}\bigl[Y_t \mathbf{1}_{\{\tau_B\geq t\}}\bigr]
  &= \mathbb{E}\Bigl[
       \mathbf{1}_{\{\tau_B\geq t\}}
       \mathbb{E}\bigl[Y_t \bigm|\mathcal{F}_{t-1}\bigr]
     \Bigr] \\
  &= \mathbb{E}\Bigl[
       \mathbf{1}_{\{\tau_B\geq t\}}J_t
     \Bigr] \\
  &= J_t \mathbb{P}(\tau_B\geq t),
\end{align*}
where in the second line we used (A2), and in the third line we used that $J_t$ is deterministic.

Thus from \eqref{eq:Es-tau-n-expansion} we obtain
\begin{equation}
  \mathbb{E}\bigl[S_{\tau_B^{(n)}}\bigr]
  =
  \sum_{t=1}^n J_t\mathbb{P}(\tau_B\geq t).
  \label{eq:Es-tau-n-with-Jt}
\end{equation}

We now lower-bound the right-hand side by using that $J_t\geq J_{\inf}>0$ for all $t$:
\begin{align*}
  \mathbb{E}\bigl[S_{\tau_B^{(n)}}\bigr]
  \geq
  J_{\inf} \sum_{t=1}^n \mathbb{P}(\tau_B\geq t)
  = J_{\inf} \mathbb{E}[\tau_B^{(n)}],
\end{align*}
because
\begin{align*}
  \sum_{t=1}^n \mathbb{P}(\tau_B\geq t)
  = \sum_{t=1}^n \mathbb{E}\bigl[\mathbf{1}_{\{\tau_B\geq t\}}\bigr]
  = \mathbb{E}\Bigl[\sum_{t=1}^n \mathbf{1}_{\{\tau_B\geq t\}}\Bigr]
  = \mathbb{E}[\tau_B^{(n)}].
\end{align*}

Note that, $S_{\tau_B^{(n)}}\leq B+M$ for all $n$, because whenever we stop (either at time $\tau_B$ or at time $n$ before reaching $B$) we cannot exceed $B+M$ by the same argument as in \eqref{eq:overshoot-inequality}. Thus
\begin{align*}
  \mathbb{E}\bigl[S_{\tau_B^{(n)}}\bigr]\leq B+M
  \quad\text{for all }n.
\end{align*}
We therefore have
\begin{align*}
  J_{\inf}\mathbb{E}[\tau_B^{(n)}]
  \leq B+M \quad\text{for all } n.
\end{align*}
Letting $n\to\infty$ and using monotone convergence $\tau_B^{(n)}\uparrow \tau_B$, we obtain
\begin{equation}\label{eq:tau-upper-crude}
  \mathbb{E}[\tau_B] \leq \frac{B+M}{J_{\inf}} < \infty.
\end{equation}
In particular, $\tau_B$ is integrable for every $B>0$.

\medskip

Now that we know $\tau_B$ is integrable, we can safely expand $S_{\tau_B}$ as an infinite sum and swap expectation and summation.

Indeed, we can write
\begin{align*}
  S_{\tau_B}
  = \sum_{t=1}^{\tau_B} Y_t
  = \sum_{t=1}^\infty Y_t\mathbf{1}_{\{\tau_B\geq t\}},
\end{align*}
where the second equality holds because only finitely many terms are non-zero (those with $t\leq\tau_B$). Taking absolute values,
\begin{align*}
  \sum_{t=1}^\infty |Y_t|\mathbf{1}_{\{\tau_B\geq t\}}
  \leq
  \sum_{t=1}^\infty M\mathbf{1}_{\{\tau_B\geq t\}}
  = M\tau_B,
\end{align*}
and $\mathbb{E}[M\tau_B]<\infty$ by \eqref{eq:tau-upper-crude}. Therefore the sum is integrable and Fubini's theorem gives
\begin{equation}
  \mathbb{E}\bigl[S_{\tau_B}\bigr]
  = \sum_{t=1}^\infty \mathbb{E}\bigl[Y_t\mathbf{1}_{\{\tau_B\geq t\}}\bigr].
  \label{eq:Es-tau-expansion}
\end{equation}

Using $\{\tau_B\geq t\}\in\mathcal{F}_{t-1}$ and (A2), we obtain
\begin{align*}
  \mathbb{E}\bigl[Y_t\mathbf{1}_{\{\tau_B\geq t\}}\bigr]
  = \mathbb{E}\Bigl[
      \mathbf{1}_{\{\tau_B\geq t\}}
      \mathbb{E}\bigl[Y_t \bigm| \mathcal{F}_{t-1}\bigr]
    \Bigr]
  = \mathbb{E}\bigl[\mathbf{1}_{\{\tau_B\geq t\}} J_t\bigr]
  = J_t \mathbb{P}(\tau_B\geq t).
\end{align*}
Therefore,
\begin{equation}
  \mathbb{E}\bigl[S_{\tau_B}\bigr]
  = \sum_{t=1}^\infty J_t\mathbb{P}(\tau_B\geq t).
  \label{eq:Es-tau-Jt-representation}
\end{equation}

Recall from \eqref{eq:overshoot-inequality} that
\begin{align*}
  B \leq S_{\tau_B} < B+M
  \quad\text{on } \{\tau_B<\infty\}.
\end{align*}
Since $\mathbb{P}(\tau_B<\infty)=1$, taking expectations gives
\begin{equation}
  B \leq \mathbb{E}\bigl[S_{\tau_B}\bigr] < B+M.
  \label{eq:Es-tau-bounds}
\end{equation}

Combining \eqref{eq:Es-tau-Jt-representation} and \eqref{eq:Es-tau-bounds} yields the key inequality
\begin{equation}
  B
  \leq
  \sum_{t=1}^\infty J_t\mathbb{P}(\tau_B\geq t)
  <
  B+M.
  \label{eq:key-ineq}
\end{equation}

Define the deviation sequence
\begin{align*}
  \Delta_t := J_t - J_\infty, \qquad t\geq 1.
\end{align*}
Then $|\Delta_t|\leq J_{\sup} + J_\infty <\infty$ for all $t$, and by assumption,
\begin{align*}
  \Delta_t \longrightarrow 0
  \quad\text{as } t\to\infty.
\end{align*}

We rewrite the sum in \eqref{eq:key-ineq} as
\begin{align}
  \sum_{t=1}^\infty J_t\mathbb{P}(\tau_B\geq t)
  &= \sum_{t=1}^\infty \bigl(J_\infty + \Delta_t\bigr)\mathbb{P}(\tau_B\geq t)
  \nonumber\\
  &= J_\infty \sum_{t=1}^\infty \mathbb{P}(\tau_B\geq t)
     + \sum_{t=1}^\infty \Delta_t \mathbb{P}(\tau_B\geq t).
  \label{eq:drift-decomposition}
\end{align}

The first sum is simply $J_\infty \mathbb{E}[\tau_B]$, because
\begin{align*}
  \sum_{t=1}^\infty \mathbb{P}(\tau_B\geq t)
  = \sum_{t=1}^\infty \mathbb{E}\bigl[\mathbf{1}_{\{\tau_B\geq t\}}\bigr]
  = \mathbb{E}\Bigl[\sum_{t=1}^\infty \mathbf{1}_{\{\tau_B\geq t\}}\Bigr]
  = \mathbb{E}[\tau_B],
\end{align*}
where the interchange of summation and expectation is justified because
\begin{align*}
  \sum_{t=1}^\infty \mathbf{1}_{\{\tau_B\geq t\}} = \tau_B
  \quad\text{and}\quad \mathbb{E}[\tau_B]<\infty.
\end{align*}

Thus \eqref{eq:drift-decomposition} becomes
\begin{equation}
  \sum_{t=1}^\infty J_t\mathbb{P}(\tau_B\geq t)
  = J_\infty\mathbb{E}[\tau_B]
    + R_B,
  \label{eq:Es-tau-decomposition}
\end{equation}
where we have defined the remainder term
\begin{align*}
  R_B := \sum_{t=1}^\infty \Delta_t\mathbb{P}(\tau_B\geq t).
\end{align*}

Plugging \eqref{eq:Es-tau-decomposition} into \eqref{eq:key-ineq}, we obtain
\begin{equation}
  B
  \leq
  J_\infty\mathbb{E}[\tau_B] + R_B
  <
  B+M.
  \label{eq:key-ineq-with-RB}
\end{equation}

\medskip


We now show that $R_B$ is negligible compared to $B$ as $B\to\infty$.

Fix an arbitrary $\varepsilon>0$. By the convergence $\Delta_t\to 0$, there exists an integer $T=T(\varepsilon)\geq 1$ such that
\begin{align*}
  |\Delta_t| \leq \varepsilon
  \quad\text{for all } t\geq T.
\end{align*}
Also define
\begin{align*}
  C := \sup_{1\leq t<T} |\Delta_t| < \infty.
\end{align*}

Split the sum defining $R_B$ into the first $T-1$ terms and the tail:
\begin{align*}
  R_B
  &= \sum_{t=1}^{T-1} \Delta_t\mathbb{P}(\tau_B\geq t)
     +
     \sum_{t=T}^\infty \Delta_t\mathbb{P}(\tau_B\geq t).
\end{align*}
We bound the absolute value of each part separately.

For the finite part,
\begin{align*}
  \biggl|\sum_{t=1}^{T-1} \Delta_t\mathbb{P}(\tau_B\geq t)\biggr|
  \leq
  \sum_{t=1}^{T-1} |\Delta_t|\mathbb{P}(\tau_B\geq t)
  \leq
  C \sum_{t=1}^{T-1} 1
  = C(T-1)
  \leq C T.
\end{align*}

For the tail $t\geq T$,
\begin{align*}
  \biggl|\sum_{t=T}^\infty \Delta_t\mathbb{P}(\tau_B\geq t)\biggr|
  \leq
  \sum_{t=T}^\infty |\Delta_t|\mathbb{P}(\tau_B\geq t)
  \leq
  \varepsilon \sum_{t=T}^\infty \mathbb{P}(\tau_B\geq t)
  \leq
  \varepsilon\mathbb{E}[\tau_B],
\end{align*}
again using $\sum_{t=1}^\infty \mathbb{P}(\tau_B\geq t) = \mathbb{E}[\tau_B]$ (and dropping the first $T-1$ terms only makes the sum smaller).

Combining both parts, we have the uniform bound
\begin{equation}
  |R_B| \leq C T + \varepsilon\mathbb{E}[\tau_B]
  \quad\text{for all } B>0.
  \label{eq:RB-bound}
\end{equation}

From \eqref{eq:key-ineq-with-RB} we have
\begin{equation}
  J_\infty\mathbb{E}[\tau_B]
  \geq B - R_B.
  \label{eq:lower-step}
\end{equation}
Using $R_B \leq |R_B|$ together with \eqref{eq:RB-bound},
\begin{align*}
  J_\infty\mathbb{E}[\tau_B]
  \geq B - C T - \varepsilon\mathbb{E}[\tau_B].
\end{align*}
Rearranging,
\begin{align*}
  (J_\infty + \varepsilon)\mathbb{E}[\tau_B]
  \geq B - C T,
\end{align*}
so
\begin{equation}
  \mathbb{E}[\tau_B]
  \geq
  \frac{B - C T}{J_\infty + \varepsilon}.
  \label{eq:tau-lower-bound}
\end{equation}

Similarly, from the upper inequality in \eqref{eq:key-ineq-with-RB},
$-R_B \leq |R_B|$, and \eqref{eq:RB-bound},
\begin{align*}
  J_\infty\mathbb{E}[\tau_B] - |R_B|
  < B+M
  \quad\Longrightarrow\quad
  J_\infty\mathbb{E}[\tau_B] - C T - \varepsilon\mathbb{E}[\tau_B]
  < B+M,
\end{align*}
so
\begin{align*}
  (J_\infty - \varepsilon)\mathbb{E}[\tau_B]
  < B + M + C T.
\end{align*}
Because $\varepsilon>0$ is arbitrary, we obtain
\begin{equation}
  \mathbb{E}[\tau_B]
  \leq
  \frac{B + M + C T}{J_\infty - \varepsilon}.
  \label{eq:tau-upper-bound}
\end{equation}

Now divide both \eqref{eq:tau-lower-bound} and \eqref{eq:tau-upper-bound} by $B$:
\begin{align*}
  \frac{\mathbb{E}[\tau_B]}{B}
  \geq
  \frac{1 - (C T)/B}{J_\infty + \varepsilon},
  \qquad
  \frac{\mathbb{E}[\tau_B]}{B}
  \leq
  \frac{1 + (M+C T)/B}{J_\infty - \varepsilon}.
\end{align*}
Letting $B\to\infty$ (so that $(C T)/B \to 0$ and $(M+C T)/B\to 0$) gives
\begin{align*}
  \liminf_{B\to\infty}
  \frac{\mathbb{E}[\tau_B]}{B}
  \geq \frac{1}{J_\infty + \varepsilon},
  \qquad
  \limsup_{B\to\infty}
  \frac{\mathbb{E}[\tau_B]}{B}
  \leq \frac{1}{J_\infty - \varepsilon}.
\end{align*}
Since $\varepsilon>0$ was arbitrary, we may let $\varepsilon\downarrow 0$ to obtain
\begin{align*}
  \liminf_{B\to\infty}
  \frac{\mathbb{E}[\tau_B]}{B}
  \geq \frac{1}{J_\infty},
  \qquad
  \limsup_{B\to\infty}
  \frac{\mathbb{E}[\tau_B]}{B}
  \leq \frac{1}{J_\infty}.
\end{align*}
Hence the limit exists and equals $1/J_\infty$:
\begin{align*}
  \lim_{B\to\infty}
  \frac{\mathbb{E}[\tau_B]}{B}
  = \frac{1}{J_\infty}.
\end{align*}

Finally, choosing $B=B_\alpha := \log(1/\alpha)$ for $\alpha\in(0,1)$ yields
\begin{align*}
  \lim_{\alpha\downarrow 0}
  \frac{\mathbb{E}[\tau_\alpha]}{\log(1/\alpha)}
  = \frac{1}{J_\infty},
\end{align*}
which completes the proof.
\end{proof}


\begin{theorem}[Dynamic hitting-time upper bound with converging lower drift]
\label{thm:dynamic-hitting-time-lower-bound-drift}
Fix a filtered probability space
$(\Omega,\mathcal{F},(\mathcal{F}_t)_{t\geq 0},\mathbb{P})$.
Let $(Y_t)_{t\geq 1}$ be a sequence of integrable random variables adapted to
$(\mathcal{F}_t)$, and define the partial sums
\begin{align*}
S_t := \sum_{i=1}^t Y_i, \qquad t \geq 1,
\end{align*}
with the convention $S_0 := 0$.

Assume the following.

\begin{enumerate}
\item[(A1)] (\emph{Bounded increments})
There exists a constant $M \in (0,\infty)$ such that
\begin{align*}
|Y_t| \leq M \quad \text{almost surely for all } t \geq 1 .
\end{align*}

\item[(A2$\ge$)] (\emph{Positive, converging lower conditional drift})
There exists a deterministic sequence $(J_t)_{t\geq 1}$ and constants
$0 < J_{\inf} \leq J_{\sup} < \infty$ such that
\begin{align*}
\mathbb{E}[Y_t \mid \mathcal{F}_{t-1}] \geq J_t
\quad \text{almost surely for all } t \geq 1 ,
\end{align*}
and
\begin{align*}
J_{\inf} \leq J_t \leq J_{\sup} \quad \text{for all } t \geq 1 ,
\end{align*}
with
\begin{align*}
J_t \longrightarrow J_\infty \in (0,\infty)
\quad \text{as } t \to \infty .
\end{align*}

\item[(A3)] (\emph{Stopping rule})
For each threshold $B > 0$, define the stopping time
\begin{align*}
\tau_B := \inf\{ t \geq 1 : S_t \geq B \},
\end{align*}
with the convention $\inf\emptyset := +\infty$.
\end{enumerate}

Then for every $B > 0$, the stopping time $\tau_B$ is integrable. Moreover,
for every $\varepsilon \in (0, J_\infty)$ there exists a finite constant
$C_\varepsilon$ such that
\begin{align*}
\mathbb{E}[\tau_B]
\leq \frac{B + M + C_\varepsilon}{J_\infty - \varepsilon}
\quad \text{for all } B > 0 .
\end{align*}
Consequently,
\begin{align*}
\limsup_{B \to \infty}
\frac{\mathbb{E}[\tau_B]}{B}
\leq \frac{1}{J_\infty} .
\end{align*}

Equivalently, if $B_\alpha := \log(1/\alpha)$ and
\begin{align*}
\tau_\alpha := \inf\{ t \geq 1 : S_t \geq B_\alpha \},
\qquad \alpha \in (0,1),
\end{align*}
then
\begin{align*}
\limsup_{\alpha \downarrow 0}
\frac{\mathbb{E}[\tau_\alpha]}{\log(1/\alpha)}
\leq \frac{1}{J_\infty} .
\end{align*}
\end{theorem}


\begin{proof}

Define the deterministic partial sums
\begin{align*}
A_t := \sum_{i=1}^t J_i, \qquad A_0 := 0,
\end{align*}
and the excess process
\begin{align*}
Z_t := S_t - A_t = \sum_{i=1}^t (Y_i - J_i), \qquad Z_0 := 0.
\end{align*}
By assumption (A2$\ge$),
\begin{align*}
\mathbb{E}[Z_t \mid \mathcal{F}_{t-1}]
= Z_{t-1} + \mathbb{E}[Y_t - J_t \mid \mathcal{F}_{t-1}]
\geq Z_{t-1}
\quad \text{a.s.},
\end{align*}
so $(Z_t)$ is a submartingale.

For $n \in \mathbb{N}$ define the bounded stopping time
\begin{align*}
\tau_B^{(n)} := \tau_B \wedge n .
\end{align*}

We claim that for every $n$,
\begin{align*}
S_{\tau_B^{(n)}} \leq B + M \quad \text{a.s.}
\end{align*}
Indeed, on $\{\tau_B \leq n\}$ we have $\tau_B^{(n)} = \tau_B$ and
$S_{\tau_B - 1} < B$ by definition of $\tau_B$, while $Y_{\tau_B} \leq M$ a.s.,
hence
\begin{align*}
S_{\tau_B} = S_{\tau_B - 1} + Y_{\tau_B} < B + M .
\end{align*}
On $\{\tau_B > n\}$ we have $\tau_B^{(n)} = n$ and $S_n < B$, so again
$S_{\tau_B^{(n)}} < B + M$.
This proves the claim.

Next, we show that
\begin{align*}
\mathbb{E}[Z_{\tau_B^{(n)}}] \geq 0,
\end{align*}
or equivalently,
$\mathbb{E}[S_{\tau_B^{(n)}}] \geq \mathbb{E}[A_{\tau_B^{(n)}}].$

Since $\tau_B^{(n)} \leq n$,
\begin{align*}
Z_{\tau_B^{(n)}}
&= \sum_{t=1}^{\tau_B^{(n)}} (Y_t - J_t)
= \sum_{t=1}^n \mathbf{1}\{\tau_B^{(n)} \geq t\} (Y_t - J_t).
\end{align*}
For $t \leq n$, $\{\tau_B^{(n)} \geq t\} = \{\tau_B \geq t\} \in \mathcal{F}_{t-1}$,
since $\tau_B$ is a stopping time. Therefore,
\begin{align*}
\mathbb{E}\!\left[\mathbf{1}\{\tau_B^{(n)} \geq t\}(Y_t - J_t)\right]
&= \mathbb{E}\!\left[
\mathbb{E}\!\left[
\mathbf{1}\{\tau_B^{(n)} \geq t\}(Y_t - J_t)
\mid \mathcal{F}_{t-1}
\right]\right] \\
&= \mathbb{E}\!\left[
\mathbf{1}\{\tau_B^{(n)} \geq t\}
\mathbb{E}[Y_t - J_t \mid \mathcal{F}_{t-1}]
\right] \\
&\geq 0 ,
\end{align*}
where the inequality follows from assumption (A2$\ge$).
Summing over $t = 1,\dots,n$ yields $\mathbb{E}[Z_{\tau_B^{(n)}}] \geq 0$.

Thus, we have that
\begin{align*}
\mathbb{E}[A_{\tau_B^{(n)}}]
\leq \mathbb{E}[S_{\tau_B^{(n)}}]
\leq B + M
\quad \text{for all } n .
\end{align*}
Since $J_t \geq J_{\inf} > 0$, the sequence $(A_t)$ is increasing and
$A_t \to \infty$ as $t \to \infty$. Because $\tau_B^{(n)} \uparrow \tau_B$,
the monotone convergence theorem yields
\begin{align*}
\mathbb{E}[A_{\tau_B}]
= \lim_{n \to \infty} \mathbb{E}[A_{\tau_B^{(n)}}]
\leq B + M .
\end{align*} 
Moreover, since
$A_{\tau_B} \geq J_{\inf} \tau_B$,
\begin{align*}
J_{\inf} \mathbb{E}[\tau_B] \leq \mathbb{E}[A_{\tau_B}] \leq B + M ,
\end{align*}
so $\tau_B$ is integrable.

Fix $\varepsilon \in (0, J_\infty)$. Since $J_t \to J_\infty$, there exists
$N = N(\varepsilon)$ such that
\begin{align*}
J_t \geq J_\infty - \varepsilon \quad \text{for all } t \geq N .
\end{align*}
Define the finite constant
\begin{align*}
C_\varepsilon
:= \sup_{0 \leq t \leq N-1} \bigl((J_\infty - \varepsilon)t - A_t\bigr).
\end{align*}
Then for all $t \geq 0$,
\begin{align*}
A_t \geq (J_\infty - \varepsilon)t - C_\varepsilon .
\end{align*}

Applying the bound from in the previous display at the random time $\tau_B$ and taking expectations yields
\begin{align*}
\mathbb{E}[A_{\tau_B}]
\geq (J_\infty - \varepsilon)\mathbb{E}[\tau_B] - C_\varepsilon .
\end{align*}
Combining this with $\mathbb{E}[A_{\tau_B}] \leq B + M$ yields
\begin{align*}
(J_\infty - \varepsilon)\mathbb{E}[\tau_B]
\leq B + M + C_\varepsilon ,
\end{align*}
and therefore
\begin{align*}
\mathbb{E}[\tau_B]
\leq \frac{B + M + C_\varepsilon}{J_\infty - \varepsilon} .
\end{align*}
Dividing by $B$ and letting $B \to \infty$, then letting $\varepsilon \downarrow 0$,
gives
\begin{align*}
\limsup_{B \to \infty} \frac{\mathbb{E}[\tau_B]}{B}
\leq \frac{1}{J_\infty} .
\end{align*}
This completes the proof.
\end{proof}
